---
title: "lab5_part2"
author: "Kelsie Fronheiser"
date: "2/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
#library(tsibble)
#library(feasts)
#library(fable)
library(here)
library(sf)
library(tmap)
```

```{r}
ca_counties <- read_sf(here("data","ca_counties","CA_Counties_TIGER2016.shp"))
```

Let’s simplify it by only keeping two attributes: NAME (county name) and ALAND (land area), then renaming those to county_name and land_area.
```{r}
ca_subset <- ca_counties %>% 
  select(NAME, ALAND) %>% 
  rename(county_name = NAME, land_area = ALAND)
```

We should notice something very important about a simple features (sf) object: it just assumes you want to keep the spatial information, and you can work with the rest of the data as if it’s a non-spatial data frame (and the spatial information just “sticks” - hence the term “sticky geometry”). So even though we only called NAME and ALAND in the select() function, we see that the geometry column still exists!

Use st_crs() to check the existing CRS for spatial data. We see that this CRS is WGS84 (epsg: 3857)
```{r}
ca_subset %>% st_crs()
```

```{r}
ggplot(data = ca_subset) +
  geom_sf(aes(fill = land_area), color = "white", size = 0.1) +
  theme_void() +
  scale_fill_gradientn(colors = c("cyan","blue","purple"))
```

## Invasive red sesbania records

```{r}
sesbania <- read_sf(here("data","red_sesbania","ds80.shp"))

# Check the CRS:
sesbania %>% st_crs()
```

Notice that this CRS is different from the California counties CRS, so we’ll want to update it to match. Use st_transform() to update the CRS:
```{r}
sesbania <- st_transform(sesbania, 3857)

# Then check it: 
sesbania %>% st_crs()
```

Cool, now they have the same CRS.

## Plotting them together
```{r}
ggplot() +
  geom_sf(data = ca_subset) +
  geom_sf(data = sesbania, size = 1, color = "red")
```

Let’s say we want to find the count of red sesbania observed locations in this dataset by county. How can I go about joining these data so that I can find counts? Don’t worry…st_join() has you covered for spatial joins!

```{r}
ca_sesbania <- ca_subset %>% 
  st_join(sesbania)
```

And then we can find counts (note: these are not counts for individual plants, but by record in the dataset) by county:
```{r}
sesbania_counts <- ca_sesbania %>% 
  count(county_name)
```
Then we can plot a choropleth using the number of records for red sesbania as the fill color (instead of what we used previously, land area):
```{r}
ggplot(data = sesbania_counts) +
  geom_sf(aes(fill = n), color = "white", size = 0.1) +
  scale_fill_gradientn(colors = c("lightgray","orange","red")) +
  theme_minimal() +
  labs(fill = "Number of S. punicea records")
```

So we see that we can still use our usual wrangling skills! Let’s do a bit more for fun, just to prove that our existing wrangling skills still work with spatial data - the spatial information just sticks to it! Only plot the county with the greatest number of red sesbania records (Solano), and make a map of those locations (yeah there are many ways to do this):

```{r}
# Subset of sesbania point locations only in Solano County
solano_sesbania <- sesbania %>% 
  filter(COUNTY == "Solano")

# Only keep Solano polygon from California County data
solano <- ca_subset %>% 
  filter(county_name == "Solano")

ggplot() +
  geom_sf(data = solano) +
  geom_sf(data = solano_sesbania)
```

## Making an interpretive map with tmap
Sometimes we’ll want to make a map interactive so that audience members can zoom in, explore different areas, etc. We can use the {tmap} package to create an interactive map. Let’s make one for our California counties (fill aesthetic by land area) with the red sesbania locations on top:
```{r}
# Set the viewing mode to "interactive":
tmap_mode(mode = "view")

# Then make a map (with the polygon fill color updated by variable 'land_area', updating the color palette to "BuGn"), then add another shape layer for the sesbania records (added as dots):
tm_shape(ca_subset) +
  tm_fill("land_area", palette = "BuGn") +
  tm_shape(sesbania) +
  tm_dots()
```

