---
title: "lab5_part1"
author: "Kelsie Fronheiser"
date: "2/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tsibble)
library(feasts)
library(fable)
library(here)
```
## Part 1
To reinforce skills for wrangling, visualizing, and forecasting with time series data, we will use data on US residential energy consumption from January 1973 - October 2017 (from the US Energy Information Administration).

Dataset: U.S. Residential Energy Consumption (Jan 1973 - Oct 2017)
Units: Trillion BTU
Source: US Energy Information Administration (data.gov)
```{r}
energy <- read_csv(here("data", "energy.csv"))
```

###Don't breeze by where it tells you the type of data the columns are (i.e character)

1. Add a new column (date) that is the current month column converted to a time series class, yearmonth
2. Convert the data frame to a tsibble, with that date column as the time index
```{r}
energy_ts <- energy %>% 
mutate(date = tsibble::yearmonth(month)) %>% 
  as_tsibble(key = NULL, index = date)

```
## Plot time series
```{r}
ggplot(data = energy_ts, aes(x = date, y = res_total)) +
  geom_line() +
  labs(y = "Residential energy consumption \n (Trillion BTU)")
```

- There is seasonality.
- There is a slight overall upward trend but stability around 2005.
- It does not have a cyclical pattern. 

## Seasonplot
```{r}
energy_ts %>% 
  gg_season(y = res_total) +
  theme_minimal() +
  labs(x = "month",
       y = "residential energy consumption (trillion BTU)")
```

- The highest residential energy usage is around December / January / February
- There is a secondary peak around July & August (that’s the repeated secondary peak we see in the original time series graph)
- We can also see that the prevalence of that second peak has been increasing over the course of the time series: in 1973 (orange) there was hardly any summer peak. In more recent years (blue/magenta) that peak is much more prominent.

```{r}
energy_ts %>% gg_subseries(res_total)
```

Our takeaway here is similar: there is clear seasonality (higher values in winter months), with an increasingly evident second peak in June/July/August. This reinforces our takeaways from the raw data and seasonplots.


##Decomposition
```{r}
# Find STL decomposition
dcmp <- energy_ts %>% 
  model(STL(res_total ~ season()))

components(dcmp)


#Visualize the decomposed components
components(dcmp) %>%
  autoplot() +
  theme_minimal()
```


##Autocorrelation

```{r}
energy_ts %>% 
  ACF(res_total) %>% 
  autoplot()
```

And yep, we see that observations separated by 12 months are the most highly correlated, reflecting strong seasonality we see in all of our other exploratory visualizations.

## Forecasting by Holt-Winters exponential smoothing

To create the model below, we specify the model type (exponential smoothing, ETS), then tell it what type of seasonality it should assume using the season("") expression, where “N” = non-seasonal (try changing it to this to see how unimpressive the forecast becomes!), “A” = additive, “M” = multiplicative. Here, we’ll say seasonality is multiplicative due to the change in variance over time and also within the secondary summer peak:
```{r}
# create the model
energy_fit <- energy_ts %>% 
  model(
    ets = ETS(res_total ~ season("M"))
  )

# Forecast using the model 10 years into the future:
energy_forecast <- energy_fit %>% 
  forecast(h = "10 years")

#Plot just the forecasted values (w/ 80 & 95% CIs):

energy_forecast %>% 
  autoplot()


```

```{r}
# Or plot it added to the original data:
energy_forecast %>% 
  autoplot(energy_ts)
```

## Assessing Residuals
We can use broom::augment() to append our original tsibble with what the model predicts the energy usage would be based on the model. Let’s do a little exploring through visualization.

First, use broom::augment() to get the predicted values & residuals:
```{r}
# Append the predicted values (and residuals) to original energy data
energy_predicted <- broom::augment(energy_fit)

# Use View(energy_predicted) to see the resulting data frame
```
Now, plot the actual energy values (res_total), and the predicted values (stored as .fitted) atop them:
```{r}
ggplot(data = energy_predicted) +
  geom_line(aes(x = date, y = res_total)) +
  geom_line(aes(x = date, y = .fitted), color = "red")
```

Cool, those look like pretty good predictions!

Now let’s explore the residuals. Remember, some important considerations: Residuals should be uncorrelated, centered at 0, and ideally normally distributed. One way we can check the distribution is with a histogram:
```{r}
ggplot(data = energy_predicted, aes(x = .resid)) +
  geom_histogram()

```

Normally distributed, centered at 0
